Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-08-30 RNARedPrint.tex
Rename main file

\PassOptionsToPackage{utf8}{inputenc}
\documentclass{bioinfo}
\copyrightyear{2018} \pubyear{2018}

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Manuscript Category}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Rolf's includegraphicstop
\makeatletter
\newsavebox{\@alignepsbox}
\newlength{\@aligneps}
\newcommand{\includegraphicstop}[2][]{%
\sbox{\@alignepsbox}{\includegraphics[#1]{#2}}%
\setlength{\@aligneps}{-\ht\@alignepsbox}%
\addtolength{\@aligneps}{2ex}%
\raisebox{\@aligneps}{\usebox{\@alignepsbox}}}
\makeatother

@@

%\newtheorem{algorithm}{Algorithm}
\newtheorem{axiom}{Axiom}
\newtheorem{hypothesis}{Working Hypothesis}
%\renewenvironment{proof}[1][]{\noindent \em Proof\ifthenelse{\equal{#1}{}}{}{ (#1)}:~}{}

@@

\begin{document}
\firstpage{1}

\subtitle{Research Article}

\title[Efficient Sampling for Multi-Target RNA Design]{Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures}
\author[Hammer, Ponty, Wang and Will]{Stefan Hammer\,$^{\text{\sfb 1,2,3}}$, Yann Ponty\,$^{\text{\sfb 4,5,}\star}$, Wei Wang\,$^{\text{\sfb 4}}$ and Sebastian Will\,$^{\text{\sfb 2}}$}
\address{$^{\text{\sf 1}}$University Leipzig, Department of Computer Science and Interdisciplinary Center for Bioinformatics, 04107 Leipzig, Germany;
$^{\text{\sf 2}}$University of Vienna, Faculty of Chemistry, Department of Theoretical Chemistry, 1090 Vienna, Austria;
$^{\text{\sf 3}}$University of Vienna, Faculty of Computer Science, Research Group Bioinformatics and
Computational Biology, 1090 Vienna, Austria;
$^{\text{\sf 4}}$CNRS UMR 7161 LIX, Ecole Polytechnique, Bat. Turing, 91120 Palaiseau, France;
and $^{\text{\sf 5}}$ AMIBio team, Inria Saclay, Bat Alan Turing, 91120 Palaiseau, France}

\corresp{$^\ast$To whom correspondence should be addressed.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\abstract{\textbf{Motivation:} The design of multi-stable RNA molecules has important applications in biology, medicine, and biotechnology. Synthetic design approaches profit strongly from effective in-silico methods, which can tremendously impact their cost and feasibility. \\
  \textbf{Results:} We revisit a central ingredient of most in-silico
  design methods: the sampling of sequences for the design of
  multi-target structures, possibly including pseudoknots. For this
  task, we present the efficient, tree decomposition-based algorithm
  \ourprog{}. Our fixed parameter tractable approach is underpinned by
  establishing the $\#${\sf P}-hardness of uniform sampling. Modeling
  the problem as a constraint network, \ourprog{} supports generic
  Boltzmann-weighted sampling for arbitrary additive RNA energy
  models; this enables the generation of RNA sequences meeting
  specific goals like expected free energies or \GCb-content. Finally,
  we empirically study general properties of the approach and generate
  biologically relevant multi-target Boltzmann-weighted designs for a
  common design benchmark, including pseudoknots. Generating seed sequences with \ourprog{}, we demonstrate significant improvements over the previously best multi-target sampling strategy.\\
  \textbf{Availability:} Our software is freely available at: \url{https://github.com/yannponty/RNARedPrint}\\
  \textbf{Contact:} \href{yann.ponty@lix.polytechnique.fr}{yann.ponty@lix.polytechnique.fr}\\
  \textbf{Supplementary information:} Supplementary data are available
  at \textit{Bioinformatics} online.}

\maketitle

\enlargethispage{35.1pt}
\section{Introduction}
\parHead{Design, applications and motivation for multiple design.}Synthetic biology endeavors the engineering of artificial biological
systems, promising broad applications in biology, biotechnology and
medicine. Centrally, this requires the design of biological
macromolecules with highly specific properties and programmable functions.
RNA present itself as a well-suited tool for a
rational design targeting specific functions~\citep{Kushwaha2016}. RNA function is tightly
coupled to the formation of secondary structure, as well as changes in
base pairing propensities and the accessibility of regions, e.g.\ by
burying or exposing interaction sites~\citep{Rodrigo2014}. At the same time, the
thermodynamics of RNA secondary structure is well understood and its prediction is
computationally tractable~\citep{McCaskill1990}. Thus,  structure can serve as effective
proxy for rational design approaches, ultimately targeting catalytic~\citep{Zhang2013} or regulatory functions.

@@

%\Final{
%Extensions:
%\begin{enumerate}
%  \item What if the input graph is not bipartite? \\
%  Extract a minimal subset of edges such that new graph is bipartite!\\
%  $\to{}$ Graph bipartization (cf implementation by Hüffner, JGAA 2009)
%  \item Is the optimization version in the \emph{weighted base-pair model} already {\sf NP}-hard?
%  \item Does it help to know that there are cliques of size $k$ during the tree decomposition?\\
%  If anything, it provides a lower bound on tree decomposition (helps branch and bound algorithms)
%  \item What if some degree of freedom is allowed in the structure definition?\\ Can we still use tree decomposition?\\
%  This looks like a very reasonable first step towards the definition of an {\em optimal} DP decomposition for RNA folding (with Pseudoknots) or RNA-RNA interactions\ldots
%  \item Structures can also be penalized, but they must be specified explicitly.\\
%  However, any competing structure must consist of helices, and it is possible to penalize each of the $\Theta(n^2)$ helices of any given length $L$. What would be the impact on treewidth?
%  \item More general energy  models $\to$ Turner. (Pseudoknots)
%  For this description, we assume that all input structures are non-crossing, while the framework would otherwise support crossing input structures as well as according energy models, e.g.{} the Hotknots model of~\citet{Ren2005}, by considering special pseudoknot loops.
%\end{enumerate}\label{lbl:theend}
%}

@@

\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
\bibliography{biblio}

\end{document}

>>>>

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%TODO uncomment
%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Rolf's includegraphicstop
\makeatletter
\newsavebox{\@alignepsbox}
\newlength{\@aligneps}
\newcommand{\includegraphicstop}[2][]{%
\sbox{\@alignepsbox}{\includegraphics[#1]{#2}}%
\setlength{\@aligneps}{-\ht\@alignepsbox}%
\addtolength{\@aligneps}{2ex}%
\raisebox{\@aligneps}{\usebox{\@alignepsbox}}}
\makeatother

@@

%\newtheorem{algorithm}{Algorithm}
\newtheorem{axiom}{Axiom}
\newtheorem{hypothesis}{Working Hypothesis}
\newenvironment{proof}[1][]{\noindent \em Proof\ifthenelse{\equal{#1}{}}{}{ (#1)}:~}{}

@@

\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%[Efficient Sampling for Multi-Target RNA Design]
\title{Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \author[Hammer, Ponty, Wang and Will]{Stefan Hammer\,$^{\text{\sfb 1,2,3}}$, Yann Ponty\,$^{\text{\sfb 4,5,}\star}$, Wei Wang\,$^{\text{\sfb 4}}$ and Sebastian Will\,$^{\text{\sfb 2}}$}

\author[
    addressref={aff1,aff2,aff3},
    email={stefan.hammer@uni-leipzig.de}
]{\inits{F}\fnm{Stefan} \snm{Hammer}}
%
\author[
    addressref={aff4,aff5},
    corref={aff4},                       % id of corresponding address, if any
    email={yann.ponty@lix.polytechnique.fr}
]{\inits{F}\fnm{Yann} \snm{Ponty}}
%
\author[
    addressref={aff4},
    email={wei.wang@lix.polytechnique.fr}
]{\inits{F}\fnm{Wei} \snm{Wang}}
%
\author[
    addressref={aff2,aff3},
    %corref={aff2},                       % id of corresponding address, if any
    email={will@tbi.univie.ac.at}
]{\inits{S}\fnm{Sebastian} \snm{Will}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \address{$^{\text{\sf 1}}$University Leipzig, Department of Computer Science and Interdisciplinary Center for Bioinformatics, 04107 Leipzig, Germany;
\address[id=aff1]{%                           % unique id
  \orgname{Dept.\ Computer Science, and Interdisciplinary Center
      for Bioinformatics, Univ.\ Leipzig}, 
  \street{H{\"a}rtelstr.\ 16-18},                     
  \postcode{D-04107},   
  \city{Leipzig}, 
  \cny{Germany}  
}

% $^{\text{\sf 2}}$University of Vienna, Faculty of Chemistry, Department of Theoretical Chemistry, 1090 Vienna, Austria;
\address[id=aff2]{%
    \orgname{Dept.\ Theoretical Chemistry, Univ.\ Vienna},
    \street{W{\"a}hringerstr.\ 17},
    \postcode{A-1090}
    \city{Wien},
    \cny{Austria}
}

% $^{\text{\sf 3}}$University of Vienna, Faculty of Computer Science, Research Group Bioinformatics and
% Computational Biology, 1090 Vienna, Austria;
\address[id=aff3]{%
    \orgname{Bioinformatics and Computational Biology Research Group, Univ.\ Vienna},
    \street{W{\"a}hringerstr.\ 17},
    \postcode{A-1090}
    \city{Wien},
    \cny{Austria}
}

% $^{\text{\sf 4}}$CNRS UMR 7161 LIX, Ecole Polytechnique, Bat. Turing, 91120 Palaiseau, France;
\address[id=aff4]{%
    \orgname{CNRS UMR 7161 LIX, Ecole Polytechnique},
    \street{Bat. Alan Turing},
    \postcode{91120}
    \city{Palaiseau},
    \cny{France}
}

% $^{\text{\sf 5}}$ AMIBio team, Inria Saclay, Bat Alan Turing, 91120 Palaiseau, France}
\address[id=aff5]{%
    \orgname{AMIBio team, Inria Saclay},
    \street{Bat. Alan Turing},
    \postcode{91120}
    \city{Palaiseau},
    \cny{France}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}
\begin{abstract}
  \parttitle{Background}
  The design of multi-stable RNA molecules has important applications in biology, medicine, and biotechnology. Synthetic design approaches profit strongly from effective in-silico methods, which can tremendously impact their cost and feasibility. \\
  \parttitle{Results} We revisit a central ingredient of most in-silico
  design methods: the sampling of sequences for the design of
  multi-target structures, possibly including pseudoknots. For this
  task, we present the efficient, tree decomposition-based algorithm
  \ourprog{}. Our fixed parameter tractable approach is underpinned by
  establishing the $\#${\sf P}-hardness of uniform sampling. Modeling
  the problem as a constraint network, \ourprog{} supports generic
  Boltzmann-weighted sampling for arbitrary additive RNA energy
  models; this enables the generation of RNA sequences meeting
  specific goals like expected free energies or \GCb-content. Finally,
  we empirically study general properties of the approach and generate
  biologically relevant multi-target Boltzmann-weighted designs for a
  common design benchmark, including pseudoknots. Generating seed sequences with \ourprog{}, we demonstrate significant improvements over the previously best multi-target sampling strategy.\\

  \smallskip
  \noindent
  Our free software is available at: \url{https://github.com/yannponty/RNARedPrint}\\
  Supplementary data are available online.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
  \kwd{RNA multi-target design}
  \kwd{RNA secondary structure}
  \kwd{Multi-dimensional Boltzmann sampling}
  \kwd{{$\#$}{\sf P}-hardness of RNA design}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
\parHead{Design, applications and motivation for multiple design.}Synthetic biology endeavors the engineering of artificial biological
systems, promising broad applications in biology, biotechnology and
medicine. Centrally, this requires the design of biological
macromolecules with highly specific properties and programmable functions.
RNA present itself as a well-suited tool for a
rational design targeting specific functions~\citep{Kushwaha2016}. RNA function is tightly
coupled to the formation of secondary structure, as well as changes in
base pairing propensities and the accessibility of regions, e.g.\ by
burying or exposing interaction sites~\citep{Rodrigo2014}. At the same time, the
thermodynamics of RNA secondary structure is well understood and its prediction is
computationally tractable~\citep{McCaskill1990}. Thus,  structure can serve as effective
proxy for rational design approaches, ultimately targeting catalytic~\citep{Zhang2013} or regulatory functions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
  All authors wrote the manuscript.
    % Text for this section \ldots
    % http://www.biomedcentral.com/submissions/editorial-policies\#authorship

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{biblio}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

 \section*{Figures}
%   \begin{figure}[h!]
%   \caption{\csentence{Sample figure title.}
%       A short description of the figure content
%       should go here.}
%       \end{figure}
% 
% \begin{figure}[h!]
%   \caption{\csentence{Sample figure title.}
%       Figure legend text.}
%       \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
% \section*{Tables}
% \begin{table}[h!]
% \caption{Sample table title. This is where the description of the table should go.}
%       \begin{tabular}{cccc}
%         \hline
%            & B1  &B2   & B3\\ \hline
%         A1 & 0.1 & 0.2 & 0.3\\
%         A2 & ... & ..  & .\\
%         A3 & ..  & .   & .\\ \hline
%       \end{tabular}
% \end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \section*{Additional Files}
%   \subsection*{Additional file 1 --- Sample additional file title}
%     Additional file descriptions text (including details of how to
%     view the file, if it is in a non-standard format or the file extension).  This might
%     refer to a multi-page table or a figure.
%
%  \subsection*{Additional file 2 --- Sample additional file title}
%    Additional file descriptions text.

\end{backmatter}
\end{document}
